---
title: "Data 605 Final"
author: "Alejandro Osborne"
date: "December 20, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(ggplot2)
library(MASS)
library(caret)
library(DT)
library(reshape)
library(corrplot)
library(Rmisc)
library(knitr)
library(dplyr)
library(psych)
library(car)
```

```{r}
htd <- read.csv("C:\\Users\\Alex O\\Downloads\\train.csv")
X <- htd$GrLivArea
Y <- htd$SalePrice
```

I chose GrLivArea: Above grade (ground) living area square feet, as the independent variable and now we create a subset with just X and Y:

```{r}
newhtd <- subset(htd, select = c(GrLivArea, SalePrice))
names(newhtd) <- c("X","Y")
```

#Probability

a.$\mathbf{P(X>x~|~Y>y)}$
```{r}
x <- quantile(X, probs = 0.25)
y <- quantile(Y, probs = 0.5)
x
y
```

```{r}
a <- length(newhtd$X[newhtd$X > x & newhtd$Y > y])/length(newhtd$Y[newhtd$Y > y])
a
```

b. $P(X>x \; \& \; Y>y)$
```{r}
b <- length(newhtd$X[newhtd$X > x & newhtd$Y>y]) / 
            nrow(newhtd)
b
```

c. $P(X<x | Y>y)$
```{r}
c <- length(newhtd$X[newhtd$X < x & newhtd$Y > y]) / 
      nrow(newhtd)
c
```

II. Does splitting the training data in this fashion make them independent? In other words, does P(XY)=P(X)P(Y)P(XY)=P(X)P(Y)? Check mathematically, and then evaluate by running a Chi Square test for association. You might have to research this.

```{r}
p_xy <- nrow(subset(htd, htd$GrLivArea > x & htd$SalePrice > y)) / nrow(htd)
p_xy
```
```{r}
p_x <- nrow(subset(htd, htd$GrLivArea > x)) / nrow(htd)
p_x
```

```{r}
p_y <- nrow(subset(htd, htd$SalePrice > y)) / nrow(htd)
p_y
```

```{r}
p_x*p_y
```
This shows that splitting it in this fashion does not them independent

Evaluate with Chi-Square Test:

```{r}
tab1 <- table(htd$GrLivArea, htd$SalePrice)
chisq.test(tab1)
```

We reject the null due to this p value - X and Y are not independent.

#Descriptive and Inferential Statistics

```{r}
summary(X)
describe(X)
hist(X)
```

```{r}
summary(Y)
describe(Y)
hist(Y)
```

```{r}
ggplot(htd, aes(x = X, y = Y)) + geom_point(color='purple') + labs(title = "Above grade (ground) living area square feet vs. Sale Price", x = "Above grade (ground) living area square feet", y = "Sale Price") + scale_y_continuous(labels = scales::comma)
```
There seems to be a strong correlation


```{r}
qplot(X,Y, data=newhtd) + stat_smooth(method=lm)
```

Box-Cox:

```{r}
summary(powerTransform(cbind(Y,X)~1, data=newhtd))
```

The estimated $\lambda$ values are both close to 0:

```{r}
lnY <- log(Y)
lnX <- log(X)
cor(lnY, lnX)
```

The correlation coefficient shows a strong relationship between the two transformed variables.

#Linear Algebra and Correlation

TotalBsmtSF: Total square feet of basement area
X1stFlrSF: First Floor square feet
LotArea: Total rooms above grade

```{r}
corrmatrix <- cor(htd[c("TotalBsmtSF","X1stFlrSF","LotArea","SalePrice")])
corrmatrix
```

```{r}
invcormat <- solve(corrmatrix)
invcormat
```  

```{r}
round(corrmatrix %*% invcormat, 15)
```

```{r}
round(invcormat %*% corrmatrix, 15)
```

#Calculus-Based Probability & Statistics

```{r}
mylndist <- fitdistr(X, "lognormal")
mylndist$estimate
mylndist$loglik
```

```{r}
set.seed(1)
sample_sel <- rlnorm(n=1000, meanlog = mylndist$estimate["meanlog"], sdlog = mylndist$estimate["sdlog"])
reald <- data.frame(GrLivingArea = X)
selected_fit <- data.frame(GrLivingArea = sample_sel)
reald$type <- "Real Data"; selected_fit$type <- "lognormal"; mytransd <- rbind(reald, selected_fit)
ggplot(mytransd, aes(x=GrLivingArea,fill = type)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity') + labs(title = "Histogram of Real Data and Simmed Lognormals")
```

It is not a perfect fit by any means but we can see that the lognormal provides a more than adequate fit to the raw data.

#Modeling

Build some type of regression model and submit your model to the competition board.  Provide your complete model summary and results with analysis.  Report your Kaggle.com user name and score.

```{r}
lregmod <- lm(SalePrice ~ LotArea + OverallQual + OverallCond + TotalBsmtSF + X1stFlrSF + GrLivArea + BsmtFullBath + FullBath + GarageCars + GarageArea, data = htd)
# summary of model
summary(lregmod)
```
```{r}
testd <- read.csv("C:\\Users\\Alex O\\Downloads\\test.csv")
```

```{r}
predictor <- predict(lregmod, testd, type="response")
datapredict <- data.frame(Id=names(predictor),SalePrice=predictor)
write.csv(datapredict,"predictor.csv", row.names=FALSE)
```

-Kaggle Score = 1.09
User Name = adcosborne
